\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-10-24 demystify.tex
updates

Fellowship is another type of funding that students can apply for (e.g., from school, industries, government). Fellowships are typically competitive and generous, and gives pretty much all benefits tuition/insurance that a TA/RA has.  Moreover, they often give higher stipend (including summer) and open doors for job opportunities (e.g., internship).  For example, a student with a Microsoft fellowship will likely get an internship at Microsoft.

@@

%\subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}

\subsubsection{Tenured or tenure-track faculty? Who do you choose?}

>>>>

Fellowship is another type of funding that students can apply for (e.g., from school, industries, government). Fellowships are typically competitive and generous, and gives pretty much all benefits tuition/insurance that a TA/RA has.  Moreover, they often give higher stipend (including summer) and open doors for job opportunities (e.g., internship).

@@

\subsubsection{Tenured, Tenure Track, Teaching,  Assistant Profs., Associate Profs}
There are several types of faculty in a university that has a CS PhD program. For example, you might hear about tenured,  tenure track, teaching faculty.  You might also see assistant, associate, full, adjunct, emeritus, teaching, research professors. Here is a primer on these titles.  Note that these are general and might not apply to all universities.

\paragraph{Tenure-line and teaching faculty} Tenure line faculty, consisting of tenured and tenured-track profs., focus on research, which includes publishing papers, obtaining grants, mentoring Ph.D students.  They often have very low teaching load (e.g., 1 per semester). \emph{Tenured} faculty are professors who have been promoted to have a permanent position (informally, very hard to fire them).   \emph{Tenure-track} faculty are (more likely young) professors who are on the track to get tenure.  Both tenure and tenured track faculty can serve as a formal advisor of Ph.D. students . This is very important because you need to have a tenure line faculty as your main PhD advisor. \S\ref{sec:tenure-vs-tenure-track} talks more about choosing between tenure and tenured-track professor as your advisor.

In contrast, \emph{teaching} (or instructional faculty or professor of practice) mainly focus on teaching. They typically teach 3-4 classes per semester and \emph{do not} have research responsibilities (e.g., they do not have to worry much about publishing papers, obtaining grants).  They can mentor Ph.D. students but they typically cannot serve as a formal advisor of Ph.D. students (i.e., they \emph{cannot graduate} PhD students).

\paragraph{Rankings} Faculty have a ranking, e.g., assistant, associate, full, regardless if they are tenure line or teaching.  Briefly, assistant means new faculty, associate means they have been promoted, and full means they are senior. A tenure track faculty starts with being an assistant professor and then gets tenure and promoted to associate (typically after 6 years) and then full (time varies, some become full within 3-4 years, some 10+ years, some remain associate).

Adjunct means they are not full-time faculty, e.g., they might be working in industry and teach a class or two.  Emeritus means they are retired but still have some affiliation with the university.  Research faculty (or scientist) are typically non-tenure line faculty who focus on research.  Due to their roles, adjunct, emeritus, and research faculty typically do not serve as a formal advisor of Ph.D. students.

\subsubsection{Tenured or tenure-track faculty? Who do you choose?}\label{sec:tenure-vs-tenure-track}
